### 2021-01-12 Analysis/Classification/mrg0138_sn_photometry.tex
adding some local txt files long un-committed * sn coordinates  - not sure if this is used anywhere * lensing_modelH.txt no longer called model H ?

#
# Preference for now for psf_flux fluxes from ePSF fitting, with uncertainty
# fluxerr.  The `bkgflux` is a local background that has already been 
# subtracted from the R=0.3" aperture `flux`.
# The magnification estimates `mu` have not been applied to the photometry
#
# time   band psf_flux apcorr flux fluxerr bkgflux zp zpsys id img mu muerr dtA dtB dtC
57588.03 F105W 0.0877 1.192 0.1311 0.0165 0.1902 23.9 ab 0 1.1 6.3 1.2 0 0 0
57588.03 F105W 2.3471 1.192 2.6767 0.0160 0.2033 23.9 ab 1 1.2 11.4 4.9 242 280 160
57588.03 F105W 0.1748 1.192 0.1922 0.0157 0.2776 23.9 ab 2 1.3 7.1 1.1 137 160 83
57587.97 F160W 0.6113 1.262 0.6702 0.0433 0.7207 23.9 ab 0 1.1 6.3 1.2 0 0 0
57587.97 F160W 3.5715 1.262 3.9274 0.0453 0.7528 23.9 ab 1 1.2 11.4 4.9 242 280 160
57587.97 F160W 1.1265 1.262 1.2476 0.0442 1.2159 23.9 ab 2 1.3 7.1 1.1 137 160 83